Searching for long-period transiting planets in the \kepler\ light curves using supervised classification

### 2015-04-09 document/ms.tex
physiscal params plots

\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\kepler}{\project{Kepler}}
\newcommand{\KT}{\project{K2}}
\newcommand{\tess}{\project{TESS}}
\newcommand{\jwst}{\project{JWST}}
\newcommand{\terra}{\project{TERRA}}
\newcommand{\pdc}{\project{PDC}}
\newcommand{\license}{MIT License}
\newcommand{\projectname}{\project{ketu}}

@@

Many of the most dynamically interesting planets will be on orbits longer
than the baseline of existing transit surveys (\kepler, \KT).
Future surveys (\tess, PLATO?) will have such a short continuous coverage that
even habitable zone planets around M-dwarfs will only present a single transit
signal.
Searches for these transiting planets are plagued by false
signals---especially when pushed to low signal-to-noise---and statistical
studies of their population are complicated by weak constraints on the
physical parameters of the system and high rates of false positives.
We develop and present a computationally expensive but tractable method of
searching for single transits using supervised classification methods from
the machine learning literature.
For each month of photometry, we train a random forest classifier on a large
number of simulated signals injected into the photometry of the same star at
different times.
Then, the model's prediction as a function of transit time is evaluated and
times above a threshold probability---chosen for each star to yield a sample
with $\sim99$~percent purity---are given as candidate transits.
Each candidate must then pass a final round of vetting including BLAH BLAH
BLAH.
Applied to every light curve in the \kepler\ archive this method yields a
list of XXXX long-period transiting planet candidates.
Using an informative prior on their eccentricities, we derive weak constraints
on the orbital periods of the candidates and demonstrate that this puts an
upper limit of ZZZZ on the occurrence rate of these planets.

@@

In \sect{est}, we extrapolate a recent model of the distribution of exoplanets
to the long periods considered in this \paper.
Using an approximate but realistic model for the detection efficiency of a
search for single transits, we estimate that $\sim 60$ single transit events
should be detectable in the archival \kepler\ dataset.
In \sect{data}, we describe the data and outline the structure exploited by
our transit search procedure.
In \sect{rfc}, we briefly describe the Random Forest classification framework
and in \sect{search} we outline the steps used to apply this model for transit
search.
This method requires the setting of many tuning parameters and the
optimization of these choices is both computationally and intellectually
challenging so in \sect{tuning} we discuss some steps in this direction.
In \sect{demo}, we demonstrate the feasibility of this method, present a
single transit discovery, and indicate some interesting weaknesses of the
method.
In \sectionname s~\sectref{params} and~\sectref{exopop}, we describe the
parameter estimation that is possible given a single transit and outline the
prospects for exoplanet population inference given a catalog of single transit
discoveries.

@@

\section{Prospects for population inference}\sectlabel{exopop}

Based on our single transit discoveries, we now want to place a limit on
the occurrence rate of long-period planets.
Since we only have one discovery thus far, we'll make the simplifying
assumption that the rate is flat in log-period and log-radius in the range
$1000\,\mathrm{d} < \period < 5000\,\mathrm{d}$ and $5\,R_\odot <
20\,R_\odot$.
For this distribution, \citet{Foreman-Mackey:2014} derived the analytic
maximum likelihood result
\begin{eqnarray}
\rate &=& \frac{N_\mathrm{obs}}{\int Q(w)\dd w}
\end{eqnarray}
where $\rate$ is the occurrence rate per natural logarithm in period and
radius, $N_\mathrm{obs}$ is the number of detections, and the integral is over
the completeness measurements for every star that has been searched.
If we make the simplifying assumption that the detectability of a transit
depends on period only through Equations~(\eqref{q-geom})
and~(\eqref{q-time}), then this integral becomes
\begin{eqnarray}
\int Q(w)\dd w &=& \sum_{k=1}^K
    \left( \frac{4\,\pi^2}{G\,M_k} \right)^{1/3} \, R_k \, T_k \, Q_k \,
    \int \period^{-8/3}\,\rp^{-1} \dd\rp\dd\period \\
 &=& \sum_{k=1}^K
    \left( \frac{4\,\pi^2}{G\,M_k} \right)^{1/3} \, R_k \, T_k \, Q_k \,
    \ln \left(\frac{\rp_\mathrm{max}}{\rp_\mathrm{min}}\right)
    \frac{3}{5} \,
    \left[{\period_\mathrm{min}}^{-5/3} - {\period_\mathrm{max}}^{-5/3}\right]
\end{eqnarray}
where the sum is over all the targeted stars.

>>>>

\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\kepler}{\project{Kepler}}
\newcommand{\KT}{\project{K2}}
\newcommand{\tess}{\project{TESS}}
\newcommand{\plato}{\project{PLATO}}
\newcommand{\jwst}{\project{JWST}}
\newcommand{\terra}{\project{TERRA}}
\newcommand{\pdc}{\project{PDC}}
\newcommand{\license}{MIT License}
\newcommand{\projectname}{\project{ketu}}

@@

Many of the most dynamically important planets have orbits longer than
the baseline of existing transit surveys (\kepler, \KT).
Future surveys (\tess, \plato) will have such a short continuous coverage that
even habitable zone planets around M-dwarfs will only present a single transit
signal.
Searches for these transiting planets are plagued by false
signals---especially when pushed to low signal-to-noise---and statistical
studies of their population are complicated by weak constraints on the
physical parameters of the system and high rates of false positives.
We develop and present a computationally expensive but tractable method of
searching for single transits using supervised classification methods from
the machine learning literature.
For each month of photometry, we train a random forest classifier on a large
number of simulated signals injected into the photometry of the same star at
different times.
Then, the model's prediction as a function of transit time is evaluated and
times above a threshold probability---chosen for each star to yield a sample
with $\sim99$~percent purity---are given as candidate transits.
Each candidate must then pass a final round of vetting including BLAH BLAH
BLAH.
Applied to every light curve in the \kepler\ archive this method yields a
list of XXXX long-period transiting planet candidates.
Using an informative prior on their eccentricities, we derive weak constraints
on the orbital periods of the candidates and demonstrate that this puts an
upper limit of ZZZZ on the occurrence rate of these planets.

@@

In \sect{est}, we extrapolate a recent model of the distribution of exoplanets
to the long periods considered in this \paper.
Using an approximate but realistic model for the detection efficiency of a
search for single transits, we estimate that $\sim 60$ single transit events
should be detectable in the archival \kepler\ dataset.
In \sect{data}, we describe the data and outline the structure exploited by
our transit search procedure.
In \sect{rfc}, we briefly describe the Random Forest classification framework
and in \sect{search} we outline the steps used to apply this model for transit
search.
This method requires the setting of many tuning parameters and the
optimization of these choices is both computationally and intellectually
challenging so in \sect{tuning} we discuss some steps in this direction.
In \sect{demo}, we demonstrate the feasibility of this method, present a
single transit discovery, and indicate some interesting weaknesses of the
method.
In \sectionname s~\sectref{params} we describe the parameter estimation that
is possible given a single transit.

% and outline the prospects for exoplanet
% population inference given a catalog of single transit discoveries.

@@

% \section{Prospects for population inference}\sectlabel{exopop}

% Based on our single transit discoveries, we now want to place a limit on
% the occurrence rate of long-period planets.
% Since we only have one discovery thus far, we'll make the simplifying
% assumption that the rate is flat in log-period and log-radius in the range
% $1000\,\mathrm{d} < \period < 5000\,\mathrm{d}$ and $5\,R_\odot <
% 20\,R_\odot$.
% For this distribution, \citet{Foreman-Mackey:2014} derived the analytic
% maximum likelihood result
% \begin{eqnarray}
% \rate &=& \frac{N_\mathrm{obs}}{\int Q(w)\dd w}
% \end{eqnarray}
% where $\rate$ is the occurrence rate per natural logarithm in period and
% radius, $N_\mathrm{obs}$ is the number of detections, and the integral is over
% the completeness measurements for every star that has been searched.
% If we make the simplifying assumption that the detectability of a transit
% depends on period only through Equations~(\eqref{q-geom})
% and~(\eqref{q-time}), then this integral becomes
% \begin{eqnarray}
% \int Q(w)\dd w &=& \sum_{k=1}^K
%     \left( \frac{4\,\pi^2}{G\,M_k} \right)^{1/3} \, R_k \, T_k \, Q_k \,
%     \int \period^{-8/3}\,\rp^{-1} \dd\rp\dd\period \\
%  &=& \sum_{k=1}^K
%     \left( \frac{4\,\pi^2}{G\,M_k} \right)^{1/3} \, R_k \, T_k \, Q_k \,
%     \ln \left(\frac{\rp_\mathrm{max}}{\rp_\mathrm{min}}\right)
%     \frac{3}{5} \,
%     \left[{\period_\mathrm{min}}^{-5/3} - {\period_\mathrm{max}}^{-5/3}\right]
% \end{eqnarray}
% where the sum is over all the targeted stars.
